---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r results='hide'}
## load the required library
library(dplyr)
library(lattice)
library(lubridate)
library(ggplot2)
library(scales)
unzip(zipfile = "activity.zip", overwrite = TRUE)
data <- read.csv(file = "activity.csv", header = TRUE)
data <- tbl_df(data)


# data <- mutate(data, hours = interval %/% 100, minutes = interval %% 100)
# data <- mutate(data, time = paste(hours, minutes, sep = ":"))
# data <- select(data, steps, date, time, interval)
# data$time <- hm(data$time)



```



## What is mean total number of steps taken per day?
```{r sumSteps_each_day}
## group the data by date to get the steps taken per day
grop_by_date <- group_by(data, date)
sumSteps_each_day <- summarize(grop_by_date, sum(steps, na.rm = TRUE))
colnames(sumSteps_each_day) <- c("date", "sumOfSteps")

hist(sumSteps_each_day$sumOfSteps, xlab = "sum of steps", main = "Histogram of steps each day", breaks = 20, col = "red")
## mean of the total steps taken per day
mean(sumSteps_each_day$sumOfSteps, na.rm = TRUE)
## median of the total steps taken per day
median(sumSteps_each_day$sumOfSteps, na.rm = TRUE)

```



## What is the average daily activity pattern?
```{r daily_activity_pattern}
## group the data by interval 

group_by_interval <- group_by(data, interval)
meanSteps_each_interval <- summarize(group_by_interval, mean(steps, na.rm = TRUE))
colnames(meanSteps_each_interval) <- c("interval", "averageSteps")

##-------------
meanSteps_each_interval <- mutate(meanSteps_each_interval, hours = interval %/% 100, minutes = interval %% 100)
meanSteps_each_interval <- mutate(meanSteps_each_interval, 
                                  time = paste(sprintf("%02d:%02d",hours, minutes), sep = ""))
meanSteps_each_interval <- select(meanSteps_each_interval, time, averageSteps, interval)
meanSteps_each_interval$time <- strptime(meanSteps_each_interval$time, format = "%H : %M")
g <- ggplot(meanSteps_each_interval, aes(time, averageSteps)) 
g + geom_line()  + scale_x_datetime(labels = date_format("%H:%M"), breaks = date_breaks("2 hour")) + labs(x = "time") + labs(y = "steps") + labs(title = "time series steps")


##-------------

## this interval has the maximum average steps
max_row <- subset(meanSteps_each_interval, averageSteps == max(averageSteps))
## get the interval with the max average steps
max_row[1, 3]
```


## Imputing missing values
```{r new_sumSteps_each_day}
## 1.calculate and report the total number of missing values in the data set
## (i use the 5-minute interval to replace)
missing_value_data <- data[is.na(data$steps),]
nrow(missing_value_data)

## 2.fill the NA values with average step in that interval
data <- mutate(data, averageSteps = rep(meanSteps_each_interval$averageSteps, (nrow(data) /nrow(meanSteps_each_interval)) ))
data[is.na(data$steps), 1] <- data[is.na(data$steps), 4]

## 3.create a new data with the missing values filled in
new_data <- select(data, steps:interval)

## 4.make a histogram
## group the new data by date
grop_by_date <- group_by(new_data, date)
sumSteps_each_day <- summarize(grop_by_date, sum(steps))
colnames(sumSteps_each_day) <- c("date", "sumOfSteps")

hist(sumSteps_each_day$sumOfSteps, xlab = "sum of steps", main = "Histogram of steps each day", breaks = 20, col = "red")

## mean of the total steps taken per day
mean(sumSteps_each_day$sumOfSteps)
## median of the total steps taken per day
median(sumSteps_each_day$sumOfSteps)

```
**From the resut, compared with the previous one, the gross level improved. Data are centered more around the mean value, beacuse we just use the mean data to replace the NA value, so the mean steps improve a lot, if the value is NA, it means we don't know the value, but it doesn't mean the real value doesn't exist. And it seems fair to us e the mean value to replace the NA value**


## Are there differences in activity patterns between weekdays and weekends?
```{r difference_in_activityPatterns}
## create a new factor with two levels
new_data$date <- ymd(new_data$date)
new_data <- mutate(new_data, weekdays = weekdays(new_data$date))
weekend <- c("Saturday", "Sunday")
## to check whether the date is in weekend
new_data <- mutate(new_data, isWeekend = is.element(new_data$weekdays, weekend))
new_data <- mutate(new_data, weekdays = factor(new_data$isWeekend, labels = c("weekday", "weekend")))
## this is new data, with the new column weekdays
new_data <- select(new_data, steps:weekdays)

## get the weekday data with the "weekdays" column is weekday
weekday_data <- subset(new_data, weekdays == "weekday")
weekday_group_by_interval <- group_by(weekday_data, interval)
weekday_meansteps_eachinterval <- summarise(weekday_group_by_interval, mean(steps))
weekday_meansteps_eachinterval <- mutate(weekday_meansteps_eachinterval, weekdays = rep("weekday", nrow(weekday_meansteps_eachinterval)))
colnames(weekday_meansteps_eachinterval) <- c("interval", "meanSteps", "weekdays")

## get the weekend data with the "weekdays" column is weekend
weekend_data <- subset(new_data, weekdays == "weekend")
weekend_group_by_interval <- group_by(weekend_data, interval)
weekend_meansteps_eachinterval <- summarise(weekend_group_by_interval, mean(steps))
weekend_meansteps_eachinterval <- mutate(weekend_meansteps_eachinterval, weekdays = rep("weekend", nrow(weekend_meansteps_eachinterval)))
colnames(weekend_meansteps_eachinterval) <- c("interval", "meanSteps", "weekdays")


## the final data reqired, with column names "interval", "meanSteps", and "weekdays"
data_final <- rbind(weekday_meansteps_eachinterval, weekend_meansteps_eachinterval)
##-----------
data_final <- mutate(data_final, hours = interval %/% 100, minutes = interval %% 100)
data_final <- mutate(data_final,
                        time = paste(sprintf("%02d:%02d",hours, minutes), sep = ""))
data_final <- select(data_final, time, meanSteps, weekdays)
data_final$time <- strptime(data_final$time, format = "%H : %M")

g <- ggplot(data_final, aes(time, meanSteps))
g + geom_line() + scale_x_datetime(labels = date_format("%H:%M"), breaks = date_breaks("2 hour")) + facet_wrap(~ weekdays, nrow = 2) + labs(x = "interval") + labs(y = "Number of steps")
##-----------


# xyplot(meanSteps ~ interval | weekdays, data = data_final, layout = c(1, 2), 
#        type = "l", ylab = "Number of steps")

```
